---
title: "p8015_hw2_sy2824"
author: "Sijia Yue"
data: "2018-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

## Data cleaning
```{r}
subway_data = 
  read.csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:vending, ada) %>%
  select(-exit_only) %>%
  mutate(entry = recode(entry,"YES" = "TRUE", "NO" = "FALSE")) %>%
  mutate(vending = recode(vending, "YES" = "TRUE", "NO" = "FALSE")) 
```

First, I read csv file into R and cleaned the format of variable names. Then, I deleted the unrelating coloums according to the question. Next, I deleted the exit_only coloum since it is totally related to the coloum "entry". Finally, I changed the "entry" and "vending" coloum from charater to logical.

## How many distinct stations?
```{r}
distinct(subway_data,line,station_name) %>%
nrow()
```
There are 465 distinct stations.


## How many stations are ADA compliant?
```{r}
filter(subway_data,ada == "TRUE") %>%
distinct(line, station_name) %>%  
nrow()
```
There are 84 ADA compliant.


## What proportion of station entrances / exits without vending allow entrance?
```{r}
num_entry_without_vending = nrow(filter(subway_data, entry == "TRUE",vending == "FALSE")) 
proportion = num_entry_without_vending/nrow(filter(subway_data,vending == "FALSE"))
proportion
```
The proportion of station entrances / exits without vending is 0.38


## Reformat data
```{r}
tidy_subway_data =
  gather(subway_data,key = route_number, value = route_served, route1:route11) %>%
  filter(route_served != 'NA') %>% 
  select(-route_number)
```

## How many distinct stations serve the A train? 

```{r}


filter(tidy_subway_data, route_served == "A") %>%
  distinct(station_name, line) %>%
  nrow()

filter(tidy_subway_data, route_served == "A") %>%
  filter(ada == "TRUE") %>%
  distinct(station_name, line) %>%
  nrow()
```
There are 60 distinct stations serve the A train. 17 of them have ADA complaint.

# Problem 2

## Data cleaning
```{r}
trash_wheel_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx",range = 'A2:N256') %>% 
  janitor::clean_names() %>%
  filter(dumpster != "NA") %>% 
  select(-month, -year) %>% 
  rename("number" = "dumpster",  "weight" = "weight_tons" , "volume" = "volume_cubic_yards")  
  
trash_wheel_data$sports_balls=as.integer(trash_wheel_data$sports_balls)
  
 
  
  



```


