---
title: "p8015_hw2_sy2824"
author: "Sijia Yue"
data: "2018-09-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(p8105.datasets)
```

# Problem 1

## Data cleaning
```{r, warning=FALSE}
subway_data = 
  read.csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:vending, ada) %>%
  select(-exit_only) %>%
  mutate(entry = recode(entry,"YES" = "TRUE", "NO" = "FALSE")) %>%
  mutate(vending = recode(vending, "YES" = "TRUE", "NO" = "FALSE")) 
```

* Read csv file into R and cleaned the format of variable names. 
* Delete the unrelating coloums according to the question. 
* Delete the exit_only coloum since it is completely related to the coloum "entry". 
* Change the "entry" and "vending" coloum from charater to logical.
* The dimisention of this dataset is `r dim(subway_data)`.
* This is not a tidy dataset because the variable `route` appears eleven times in the varibles row, this problem should be fixed.

**How many distinct stations?**
```{r}
distinct(subway_data,line,station_name) %>%
nrow()
```
There are 465 distinct stations.


**How many stations are ADA compliant?**
```{r}
filter(subway_data,ada == "TRUE") %>%
distinct(line, station_name) %>%  
nrow()
```
There are 84 ADA compliant.


**What proportion of station entrances / exits without vending allow entrance?**
```{r}
num_entry_without_vending = nrow(filter(subway_data, entry == "TRUE",vending == "FALSE")) 
proportion = num_entry_without_vending/nrow(filter(subway_data,vending == "FALSE"))
proportion
```
The proportion of station entrances / exits without vending is 0.377


## Reformat data
Since the `route` variable has eleven corresponding variables on the first row, I would use the function `gather()` to gather the route information into a new variable called`route_number`.

```{r, warning=FALSE, message=FALSE}
tidy_subway_data =
  gather(subway_data,key = route_number, value = route_served, route1:route11) %>%
  filter(route_served != 'NA') %>% 
  select(-route_number)
```

**How many distinct stations serve the A train? How many are ADA compliant?**

```{r}
filter(tidy_subway_data, route_served == "A") %>%
  distinct(station_name, line) %>%
  nrow()

filter(tidy_subway_data, route_served == "A") %>%
  filter(ada == "TRUE") %>%
  distinct(station_name, line) %>%
  nrow()
```
There are 60 distinct stations serve the A train. 17 of them have ADA complaint.

# Problem 2

## Mr. Trash Wheel dataset
* Import Mr. Trash Wheel data from the `.xlsx` file into R, in the range of coloum A to N.
* Clean the variable name format and rename `dumpster`, `weight_tons` and `volume_cubic_yards`.
* Filter the coloums that `number` is not null.
* Rounds the number of sports balls to the nearest integer and converts the result to an integer variable.

```{r}
trash_wheel_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  rename("number" = "dumpster",  "weight" = "weight_tons" , "volume" = "volume_cubic_yards")  %>% 
  filter(number != "NA")  
  
trash_wheel_data$sports_balls = as.integer(trash_wheel_data$sports_balls)
```

## Precipitation dataset
* Read data set of 2016 and 2017 precipitation data.
* Clean the variable name format.
* Filter the `month` that is not null.
* Add a new variable `year`.
* Combine the two dataset together.

```{r}
precipitation_2016_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation", range = 'A2:B15') %>%
  janitor::clean_names() %>% 
  filter(month != "NA") %>% 
  mutate(year = 2016) 
  
precipitation_2017_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = 'A2:B15') %>%
  janitor::clean_names() %>% 
  filter(month != "NA") %>% 
  mutate(year = 2017) 
    
precipitation_data = 
  bind_rows(precipitation_2016_data, precipitation_2017_data)  %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, total)
```

## Comment
In Mr. Trash Wheel dataset, there are `r nrow(trash_wheel_data)` observations and `r ncol(trash_wheel_data)` variables. The key variables are `number`, `data`, `weight` and `homes_powered`.
In precipitation dataset, there are `r nrow(precipitation_data)` observations and `r ncol(precipitation_data)` variables. The key variables are `year`, `month` and `total`.

**What was the total precipitation in 2017? **
```{r}
total_2017_data = filter(precipitation_data, year == "2017", total != "NA")
total_2017 = sum(total_2017_data$total)
```
The total precipitation in 2017 is `r total_2017` tons.

**What was the median number of sports balls in a dumpster in 2016?**
```{r}
filter(trash_wheel_data, year == "2016", sports_balls != "NA")
sports_balls_2016_median = median(trash_wheel_data$sports_balls)
```
The median number of sports balls in a dumpster in 2016 is `r sports_balls_2016_median`.

# Problem 3
* Set up a dataframe called `tidy_brfss`.
* Read data from the library.
* Clean variable name formats.
* Focus on the topic `Overall Health`.
* Select the coloums according to the question.
```{r}
tidy_brfss = 
  brfss_smart2010 %>%
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))
```
* Structure data so that values for Response (“Excellent” to “Poor”).
* Create a new variable `proportion` showing the proportion of responses that were “Excellent” or “Very Good”.
```{r}
structured_brfss = 
  spread(tidy_brfss, key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(sum = excellent + fair + good + poor + very_good) %>% 
  mutate(proportion = (excellent + very_good)/sum) 
```

**How many unique locations are included in the dataset? Is every state represented? What state is observed the most?**
```{r}
distinct(structured_brfss,locationabbr, locationdesc) %>% 
  nrow()

distinct(structured_brfss, locationabbr) %>% 
  nrow()

count(structured_brfss,locationabbr) %>% 
  arrange(desc(n))
```
There are 404 unique locations in this dataset, and every state is represented. New Jersey is observed the most.

**In 2002, what is the median of the “Excellent” response value?**
```{r}
structured_brfss_2002 = filter(structured_brfss, year == "2002") 
median(structured_brfss_2002$excellent, na.rm = TRUE)
```
The median of the "Excellent" response is 23.6

**Make a histogram of “Excellent” response values in the year 2002.**
```{r warning=FALSE, message=FALSE}
ggplot(structured_brfss_2002, aes(x = excellent)) +
  geom_histogram(binwidth = 1.5,na.rm = T) +
  labs(
    title = "Excellent response values histogram",
    x = "Excellent response values in the year of 2002",
    y = "Number"
  )
```

**Make a scatterplot showing the proportion of “Excellent” response values**
```{r warning=FALSE, message=FALSE}
filter(structured_brfss, 
       locationdesc == "NY - New York County" | locationdesc == "NY - Queens County") %>% 
ggplot(aes(x = year, y = excellent/sum, color = locationdesc)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Excellent proportion scatterplot",
    x = "Year",
    y = "Proportion of Excellent responses values"
  ) +
  theme(legend.position = "bottom")
```

